---
title: "Code for creating maps using the leaflet R package"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Install packages

```{r }
library(leaflet)
library(htmltools)
library(htmlwidgets)
```

###Map 1 (Cities visited around the world ) ###

```{r}
# Read in data
cities_visited_world <- read.csv("data/cities_visited_world.csv")

# First create color palette so that cities are colored differently depending on the continent
# "forestgreen"= #228B22
# "slateblue"= #6A5ACD
# "red"= #FF0000
# "orange"= #FFA500
# "steelblue"= #4682B4
                   
pal <- colorFactor(palette = c("#228B22", "#6A5ACD", "#FF0000", "#FFA500", "#4682B4"), 
                   domain = cities_visited_world$continent)

# Customize labels for each data point
labels = paste(cities_visited_world$city, "<br/>", cities_visited_world$country) %>% lapply(htmltools::HTML)

# Create map
cities_visited_world_map <- leaflet(cities_visited_world) %>%
  addProviderTiles("OpenStreetMap") %>%
  addCircleMarkers(lng=~lon, lat=~lat, 
                   popup = ~city, 
                   label=labels,
                   color = ~pal(continent), 
                   radius=3, 
                   stroke=TRUE, fillOpacity=0.5)
cities_visited_world_map
```

Sometimes when there are many data points, it's good to create clusters. You can do this using the clusterOptions as shown here:

```{r}
# Create map with cluster circles
cities_visited_world_map_cluster <- leaflet(cities_visited_world) %>%
  addTiles() %>%
  addCircleMarkers(lng=~lon, lat=~lat, 
                   popup = ~city, #City label from the dataset shows when clicking on a city
                   label=labels, #Customized label shows when mouse is over the city (but not clicked)
                   color = ~pal(continent), 
                   radius=3, 
                   stroke=TRUE, fillOpacity=0.5, 
                   clusterOptions = markerClusterOptions()) #Add clusters
cities_visited_world_map_cluster

```

Finally, you can save the map as a widget that you can upload to a website. The following code will create an .html file in your current directory

```{r eval=FALSE}
saveWidget(cities_visited_world_map, file="cities_visited_world_map.html")
```

### MAP 2 (USA States visited) ###

```{r}
# Read in data
states_visited_usa <- read.csv("data/states_visited_usa.csv")
```

Get polygons for US States. There are multiple ways to do this. For example:

```{r results="hide", message=FALSE}

# 1 - Using the tigris package
library(tigris)
us_states <- states(cb=TRUE, resolution="20m") 

# 2 - From a shape file downloaded from census.gov and reading it using the rgdal package
# See: https://rstudio.github.io/leaflet/shapes.html
# From https://www.census.gov/geo/maps-data/data/cbf/cbf_state.html
library(rgdal)
us_states <- readOGR("./shp/cb_2017_us_state_20m.shp",
                  layer = "cb_2017_us_state_20m", GDAL1_integer64_policy = TRUE)
```

Once you have the shape file with the US states, merge this file with the states_visited_usa data frame that shows whether you visited or not a state (1=Visited, 0=Haven't visited)

```{r}

# Merge US states polygon data with data frame indicating the states visited/not visited
states_merged <- geo_join(us_states, states_visited_usa, "NAME", "state") #geo_join is from the tigris package

# Choose colors for states visited / not visited
visited_color_1 <- "#0000CD" #Blue-ish color
not_visited_color_0 <- "#66CD00" #Green-ish color

# Other colors used for the map:
# #000000 BLACK
# #BEBEBE GRAY

# Indicate which colors will be used for the states visited / not visited
mypalette <- colorNumeric(c(not_visited_color_0, visited_color_1), c(0,1), na.color="transparent") # (or convert the $visited column to factor and use colorFactor instead of colorNumeric)
#mypalette(c(0,1)) #Colors chosen for values 0 and 1

# Create map
us_states_visited_map <- leaflet(states_merged) %>% 
  addTiles() %>% 
  setView(lat=39.83, lng=-98.58, zoom=4) %>%
  addPolygons(weight=1, smoothFactor=0.2, color = "#BEBEBE", fill=FALSE) %>% 
  addPolygons(weight=0.5, 
              smoothFactor = 0.5,
              color=~mypalette(visited),
              fillColor = ~mypalette(visited), 
              opacity = 0.5,
              fillOpacity = 0.4, 
              stroke=TRUE, 
              label = ~state, #Show state label when mouse is on the state
              highlightOptions = highlightOptions(color="#000000", weight=3, bringToFront=TRUE)) %>% #highlights moused-over polygon
  addLegend(data=states_merged, colors=c("#66CD00", "#0000CD"), labels=c("Have not visited", "Visited"), opacity=0.4, title="States visited", position="bottomleft")
us_states_visited_map

```
Save the map as an .html widget that you can use to upload to a website

```{r eval=FALSE}
saveWidget(us_states_visited_map, file="cities_visited_world_map.html")
```

For more information, see [Leaflet for R](https://rstudio.github.io/leaflet/)



